Quantifying the manageability of pollination networks in an invasion context

### 2017-02-28 paper/draft_2.Rmd
Update discussion
Search for (FDO XXX) to find sections that might require attention

```{r fig-dependency-fs, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 3.245, fig.height = 3.2, fig.cap = "**Relationships between _f~S~_ and _f~D~_ and species structural properties**. (a) In all networks were they were present, invasive species were classified as superior (_f~S~_) and driver (_f~D~_) nodes in all possible control configurations. (b) Species strength (the sum of the dependences of other species on the species of interest) is the single most important factor explaining _f~S~_. Visitation strength and degree also had an important albeit comparatively smaller effect (dashed lines correspond to one standard deviation of these factors). Invasive species are depicted as circles."}
data <- mo 
trans <- function(x, orig){
  x * attr(orig, "scaled:scale") + attr(orig, "scaled:center")
}

@@

coef_tables %<>%
	plyr::llply(function(x){
		x %>%
  magrittr::set_names(c(" ", "est.", "var", "mod.", "imp.", "C.I.")) %>%
			dplyr::select(-var, -mod.)
	})

attr(coef_tables, "subheadings") <- paste0("$", names(mod), "$")
coef_tables %>%
	xtable::xtableList(caption = "\\textbf{Factors determining species importance}. Factor estimates correspond to the average over all models that accounted for 95\\% of the AIC evidence. Confidence intervals correspond to $\\alpha = 0.05$.",
										 digits =  c(0, 0, 2, 2, 2),
										 display = c("d", "s", "f", "f", "G")) %>%
	print(booktabs = T, include.rownames = F,  caption.placement = "top", size = "\\small",comment = F, sanitize.text.function = function(x){x})
```

@@

Contrary to our initial hypothesis, we found some evidence that invaded communities might be easier to manage than uninvaded ones. 
Our results reveal, however, that this effect is comparatively small and the structural differences among different networks are more importantly related to potential differences in our ability to alter the state of the community.
Despite the small effect of invasion status at a network level, we found that invasive mutualists occupy a particularly dominant role in their communities. 
Not only changes on their abundance have the potential to propagate broadly trough the community and in turn, affect the abundances of many other species, but they are also indispensable to fully control the plant-pollinator network. 
At a community level, we demonstrate that the manageability of mutualistic networks is strongly governed by the asymmetric nature of mutual dependences—which constitute the foundations of the structure and stability of mutualistic networks [@Memmott2004; @Vazquez2004; @Bascompte2006; @Lever2014; @Astegiano2015]. 
Moreover, these mutual dependences seem to be importantly constrained by the concerted effects of both, the patterns of species richness at each trophic guild, and the species' degree distribution [@Melian2002; @Bluthgen2007]. 
Indeed, the difference between the driver-node density (*n~D~*) of our empirical networks and that of randomisations depended strongly on the evaluated null model. 
While the empirical *n~D~* was indistinguishable from that of networks with a random structure that maintained the degree or strength of each species in the community, it was much larger than that of randomisations in which the directed network was unchanged but where the observed patterns of dependence were broken.

Invasive species have been previously found to exacerbate the asymmetries in their communities [@Aizen2008; @Bartomeus2008; @Henriksson2016]. 
Although this might cause differences both at the community and the species level, we found that invasive plants are not inherently different to their native counterparts [@Stouffer2014].
Invasive plants, just like any other mutualist in our data set, tend to be classified as a superior node with a frequency (*f~S~*) proportional to the degree to which their interaction partners are collectively more dependent on them than the other way around.
Previous studies have found that super-generalists, like invasive species, play a central role in their networks [@Vila2009; @Palacio2016].
Our results take this one step further and indicate that dependence strength, rather than generalism or other metrics of centrality, is the factor that best explains the cascading effects a species could trigger on its community.

Because of the ability that our approach has to infer the magnitude of the effects that each species has on others in the community, it is tantalising to use it to select promising candidates for management interventions. 
The two indices we employed to characterise a species provide two complementary pieces of information. 
Our first index, the frequency with which a species is classified as a driver node (*f~D~*) provides an indication of the likelihood the species is part of the set of species that need to be intervened in order to control of *all* species in the community. 
The driver-node concept has received considerable attention in the structural control literature and indeed shows substantial potential to provide useful ecological insight. 
Nevertheless, we anticipate two caveats that hinder its direct utility for management applications.
First, unlike other types of complex systems, fully controlling an ecological community is realistically out of reach for all but the most simple ones; either because of the number of required interventions or the difficulties of their implementation [@Motter2015]. 
For instance, our results suggest that full control of the pollination network would require direct interventions on 40-90% of the species.
Second, @Ruths2014 established that driver nodes arise due to distinct mechanisms and therefore species with markedly different network metrics can be driver nodes in their community (Supporting information S2).

Our second index *f~S~* is, however, directly related to the likelihood a species has to affect the abundance of another species in all of the considered control strategies. 
Importantly, this is irrespective of whether controlling the entire network is desired/feasible or not. 
In fact, because superior nodes are invariantly at the beginning of a matching link, species with a high *f~S~* are more likely to be the subjects of management interventions when controlling the abundances of a target set of species—and not the entire network—is desired [@Gao2014]. 
An important advantage is that the target set of species does not has to be the same set to which interventions are applied. 
For instance, despite inconsistent outcomes in practice [@Suding2004; @Rodewald2015; @Smith2016], our results suggest that current restoration approaches that focus on direct eradication of invasive species might indeed be an effective way to modify ecosystem state. 
Nevertheless, our results also indicate that removals must be exercised with caution. 
Not only it is hard to predict the direction in which the system will change, but also, invaded communities tend to be highly dependent on invaders and therefore acutely vulnerable to their eradication [@Traveset2013; @Albrecht2014].

Although *f~S~*, in particular, is promising for identifying species priorities, our approach is different to previous attempts to quantify species importance. 
Existing metrics usually harness species features like centrality, position, co-extinction or uniqueness, to infer their effect on other species [@Allesina2004; @Jordan2006; @Jordan2009; @Lai2012]. 
Contrastingly, our control-based approach, by definition, tackles that question directly.
Although relatively simple to calculate, it has been shown that classic species-level network metrics do not necessarily reveal the best set of species to manage [@Eklof2013; @McDonald-Madden2016].
Our approach, however, is not based on a single structural metric but acknowledges the existence of multiple management strategies; some more optimal than others in specific contexts. 
Our, and other flexible approaches that take a network-wide approach might prove more useful to guide ecosystem management if we want to go beyond using network metrics to minimise topological species loss [@McDonald-Madden2016].

Despite its conceptual advantages, our approach still relies on knowledge of the network of interaction between species. 
We do not argue that management decisions should be based on a single technique. 
Nevertheless, our results and previous basic research show that as long as the proportion of sampled links is enough to provide indication of the actual degree distribution (and in turn the species dependences), *n~D~*, *f~S~* and *f~D~* can still provide a simple, straightforward, and theoretically-informed indication of the degree to which the community is self-regulated and therefore how difficult it might be to modify its state in one way or another using some of the species that integrate it [Supporting Information S5; @Nepusz2012].

In this study, we illustrate how a control-theoretic approach can be adopted in network ecology to evaluate the effect of invasions and another kind of perturbations. 
Although, because of the degree constraints imposed by bipartite networks, our pollination specific results might not be relevant to other ecological systems, the approach we propose is applicable wherever species abundances are influenced by their interactions. 
Exciting open questions lie ahead: how to design the precise "control signals" to reach a desired ecosystem state or conservation outcome? 
What are the implications of assuming linear species dynamics?
How important is to include several interaction types for our understanding of manageability and species importance?
What are the implications for species coexistence? 
Which are the trade-offs between persistence at the species and the community level?

Answering these questions might require its evaluation on different ecological systems, an explicit integration of control theory with numerical models of species densities [@Cornelius2013; @Gibson2016], and experimental tests on simple communities. 
Nevertheless, the rewards are encouraging. 
Not only from a purely ecological perspective, but also from a conservation perspective, where an integrated approach can shift our focus beyond the identification of ideal targets for intervention to design informed interventions that legitimately achieve restoration goals.

@@

Matching weight
  ~ The sum of the weights of all matched links in a particular matching.

Maximal cardinality matching
  ~ A matching with the largest possible matching size. In unweighted networks, all maximal cardinality matchings are also maximum matchings.

Maximum matching
  ~ A matching with the largest possible matching size *and* matching weight.

Superior node
  ~ The node at the start of a matched link. From the control perspective, superior nodes make up the chain(s) that propagate the control signal(s) through the network.

>>>>

```{r fig-dependency-fs, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 3.245, fig.height = 3.2, fig.cap = "**Relationships between _f~S~_ and _f~D~_ and species structural properties**. (a) In all networks were they were present, invasive species were classified as superior (_f~S~_) and driver (_f~D~_) nodes in all possible control configurations. (b) Species strength (the sum of the dependences of other species on the species of interest) is the single most important factor explaining _f~S~_. Visitation strength and degree also had an important albeit comparatively smaller effect (dashed lines correspond to $\\pm$ one standard deviation of these factors). Invasive species are depicted as circles."}
data <- mo 
trans <- function(x, orig){
  x * attr(orig, "scaled:scale") + attr(orig, "scaled:center")
}

@@

coef_tables %<>%
	plyr::llply(function(x){
		x %>%
			dplyr::select(factor, imp., est., var, mod., C.I.) %>%
			magrittr::set_names(c(" ", "imp.", "est.", "var", "mod.",  "C.I.")) %>%
			dplyr::select(-var, -mod.)
	})

attr(coef_tables, "subheadings") <- paste0("$", names(mod), "$")
coef_tables %>%
	xtable::xtableList(caption = "\\textbf{Factors explaining species importance}. Factor estimates correspond to the average over all models that accounted for 95\\% of the AIC evidence. Confidence intervals correspond to $\\alpha = 0.05$.",
										 digits =  c(0, 0, 2, 2, 2),
										 display = c("d", "s", "f", "f", "G")) %>%
	print(booktabs = T, include.rownames = F,  caption.placement = "top", size = "\\small",comment = F, sanitize.text.function = function(x){x})
```

@@

Contrary to our initial hypothesis, we found some evidence that invaded communities might be easier to manage than uninvaded ones from a control-theoretic perspective. 
Our results reveal, however, that this effect is comparatively small and the structural differences among different networks are more strongly related to potential differences in our ability to alter the state of the community via external interventions.
Despite the small effect of invasion status at a network level, we found that invasive mutualists occupy a particularly dominant role in their communities for two reasons. 
(FDO revision) First, as species with a high *f~S~*, changes on their abundance have the potential to propagate broadly through the community and in turn, affect the abundances of many other species. 
Second, as species with a high *f~D~*, they are also indispensable to fully control the plant-pollinator network. 
At a community level, we demonstrate that the manageability of mutualistic networks is strongly governed by the asymmetric nature of mutual dependences—which constitute the foundations of the structure and stability of mutualistic networks [@Memmott2004; @Vazquez2004; @Bascompte2006; @Lever2014; @Astegiano2015]. 
Moreover, these mutual dependences seem to be constrained by the effects of both the patterns of species richness at each trophic guild and a network's degree distribution [@Melian2002; @Bluthgen2007]. 
Indeed, the difference between the driver-node density (*n~D~*) of our empirical networks and that of randomisations depended strongly on the null model's randomisation approach. 
While the empirical *n~D~* was indistinguishable from that of networks with a random structure that maintained the degree of each species in the community, it was larger than that of randomisations in which the directed network was unchanged but where the observed patterns of dependence were broken.

Invasive species have been previously found to exacerbate the asymmetries in their communities [@Aizen2008; @Bartomeus2008; @Henriksson2016]. 
Although this might cause differences both at the community and the species level, we found that invasive plants are not inherently different to their native counterparts [@Stouffer2014].
Invasive plants, just like any other mutualist in our data set, tend to be classified as a superior node with a frequency (*f~S~*) proportional to the degree to which their interaction partners are collectively more dependent on them than the other way around.
Previous studies have found that supergeneralists, like invasive species, play a central role in their networks [@Vila2009; @Palacio2016].
Our results take this one step further and indicate that dependence strength, rather than generalism or other metrics of centrality, is the factor that best explains the cascading effects a species could trigger on its community.

Because of the ability that our approach has to infer the magnitude of the effects that each species has on others in the community, it is tantalising to use it to select promising candidates for management interventions. 
To this end, the two indices we employed to characterise a species provide two complementary pieces of information. 
(FDO revise grammar) Our first index, the frequency with which a species is classified as a driver node (*f~D~*) provides an indication of the likelihood that a species is part of the set of species that need to be manipulated in order to control of *all* species in the community. 
The driver-node concept has received considerable attention in the structural-control literature and indeed shows substantial potential to provide useful ecological insight. 
Nevertheless, we anticipate two caveats that hinder its direct utility for management applications.
First, unlike some other types of complex systems, fully controlling an ecological community is almost certainly out of reach for all but the simplest ones, due to either the number of required interventions or the difficulties of their implementation [@Motter2015]. 
For instance, our results suggest that full control of the pollination networks would require direct interventions on 40-90% of the species.
Second, @Ruths2014 established that driver nodes arise due to distinct mechanisms and therefore species with markedly different network metrics can be driver nodes in their community (Supporting information S2).

Our second index *f~S~* is, however, directly related to the likelihood a species affects the abundance of another species in all of the control strategies considered. 
Importantly, this is irrespective of whether controlling the entire network is desired/feasible or not. 
In fact, because superior nodes are invariably at the beginning of a matched link, species with a high *f~S~* are more likely to be the subjects of management interventions when controlling the abundances of a target set of species—as opposed to the entire network—is desired [@Gao2014]. 
An important advantage here is that the target set of species does not have to be the same set to which interventions are applied. 
For instance, despite inconsistent outcomes in practice [@Suding2004; @Rodewald2015; @Smith2016], our results suggest that current restoration approaches that focus on direct eradication of invasive species might indeed be an effective way to modify ecosystem state. 
Nevertheless, our results also indicate that removals must be exercised with caution. 
Not only it is hard to predict the direction in which the system will change, but invaded communities also tend to be highly dependent on invaders and therefore acutely vulnerable to their eradication [@Traveset2013; @Albrecht2014].

(FDO check) Despite the apparent similarities, our approach is different to previous attempts to quantify species importance. 
Existing metrics usually harness species features like centrality, position, co-extinction or uniqueness, to infer their effect on other species [@Allesina2004; @Jordan2006; @Jordan2009; @Lai2012]. 
In contrast, our control-based approach tackles that question directly.
Although they are relatively simple to calculate, it has been shown that classic species-level network metrics do not necessarily reveal the best set of species to manage [@Eklof2013; @McDonald-Madden2016].
Our approach, however, is not based on a single structural metric but instead acknowledges the existence of multiple management strategies.
(FDO check) By allowing for the fact that some are more optimal than others in different contexts, control theory implicitly highlights that management decisions should be based on a single technique. 
As such, ours and other flexible approaches that take a network-wide approach might prove more useful to guide ecosystem management if we want to go beyond using network metrics to minimise topological species loss [@McDonald-Madden2016].

Despite its conceptual advantages, our approach still relies on knowledge of the network of interaction between species. 
Nevertheless, our results and previous basic research show that, as long as the proportion of sampled links is enough to provide indication of the actual degree distribution (and in turn the species dependences), *n~D~*, *f~S~* and *f~D~* can still provide a simple, straightforward, and theoretically informed indication of the degree to which the community is self regulated and therefore how difficult it might be to modify its state in one way or another using some of its constituent species [Supporting Information S5; @Nepusz2012].

In this study, we illustrate how a control-theoretic approach can be better adopted in network ecology to evaluate the effect of invasions and other kinds of perturbations. 
Although our pollination specific results might not be directly translatable to other ecological systems because of the degree constrains imposed by bipartite networks, the approach we propose is applicable wherever species abundances are influenced by their interactions. 
Exciting open questions lie ahead. How to design the precise "control signals" to reach a desired ecosystem state or conservation outcome? 
What are the implications of assuming linear species dynamics?
How important it is to include several interaction types for our understanding of manageability and species importance?
What are the implications for species coexistence? 
Which are the trade-offs between persistence at the species and the community level? 
Answering each of these questions might require its evaluation in different ecological systems, an explicit integration of control theory with numerical models of species densities [@Cornelius2013; @Gibson2016], and experimental tests on simple communities. 
Nevertheless, the potential rewards are encouraging from both an ecological perspective as well as a conservation perspective, where an integrated approach can shift our focus beyond the identification of ideal targets for intervention to the design of informed interventions that legitimately achieve restoration goals.

@@

Matching weight
  ~ The sum of the weights of all matched links in a matching.

Maximal cardinality matching
  ~ A matching with the largest possible matching size. In unweighted/binary networks, all maximal cardinality matchings are also maximum matchings.

Maximum matching
  ~ A matching with the largest possible matching size *and* largest possible matching weight.

Superior node
  ~ The node at the start of a matched link. From the control perspective, superior nodes make up the chains that propagate the control signals through the network.
